The Emotion--Facial expression link: Evidence from Human and Automatic Expression Recognition

### 2019-11-04 manuscript_apa.Rmd
text before modif

keywords          : "Facial expression, self-report, human observer, automatic recognition."
wordcount         : "7920"

@@

header-includes:
  - \usepackage{booktabs}
  - \usepackage{float}
  - \usepackage{tabu}
  - \usepackage{wrapfig}
  - \usepackage[none]{hyphenat}
  - \floatplacement{figure}{H}
  - \usepackage{flushend}
  - \usepackage{biblatex}
---

@@

With the development of commercial automatic facial expression recognition tools [@dupre2018accuracy], industries and governments are gradually implementing this technology in order to track humans’ emotions in various scenarios (*e.g.*, in marketing, healthcare, and the automotive industry to name a few). This technology rests on the premise that facial expressions provide a direct access to individuals’ subjective feelings, and that one can read the emotions displayed on a person’s face as in an open book. This premise is derived from the Basic Emotion View which Paul Ekman is one of the famous proponents of this perspective [@ekman1987universals; @ekman1988universality; @ekman1992argument; @ekman2007directed]. Even if this premise is central to the modern mainstream approach of human emotion, recent research in affective science is challenging it. Once the Basic Emotion View is briefly described, its foremost criticisms are synthetically exposed. Inconsistencies and unpredicted findings research has uncovered come both from field observations and laboratory experiments. They relate to two different issues. The first issue is the one of the sender’s production of spontaneous facial displays. The second issue has to do with the viewer’s interpretation of these facial displays. Building on this evidence, and as a further attempt to shed light on the emotion–face link, the present study is finally exposed. Its main goal is to tackle these two issues at the same time. It aims at identifying the extent to which ordinary people’s experienced emotions are displayed through identifiable emotional facial expressions (EFE) recognized as such both by human individuals and by automatic facial expression recognition tools.

@@

```{r self_report_method}
self_report_gender <- self_report_data %>% 
  dplyr::group_by(genre_c) %>% 
  dplyr::summarise(n = n())

@@

```{r human_recognition_method}
unique_ppt_count <- human_recognition_data %>% 
  dplyr::group_by(source,SEXE_Juge, C_Juge) %>% 
  dplyr::summarise(n = n())

@@

```{r sr_hr}
comparison_sr_hr <- dplyr::inner_join(self_report_score,human_recognition_score, by = "C_Video") %>% 
  dplyr::ungroup() %>% 
  dplyr::filter(hr_emotion != "undetermined") %>% 
  dplyr::filter(sr_emotion != "undetermined") %>% 
  dplyr::mutate_if(is.character,as.factor)

@@

The overall correlation of recognition and non-recognition between self-reported emotions and human observers recognition is significant but low (`r corr_sr_hr$full_result`). In order to identify differences according to the emotional labels, encoders’ subjective feelings are compared with human observers’ recognition in a confusion matrix (Figure \ref{fig:confusionMatrix_sr_hr}).

```{r confusionMatrix_sr_hr, fig.height=6, fig.cap="\\label{fig:confusionMatrix_sr_hr}Confusion matrix between the emotion self-reported as being characteristic of the elicitation with the emotion recognized by the human observers."}
confusionMatrix_sr_hr_freq <- confusionMatrix_sr_hr$table %>% 
  as.data.frame() %>% 
  dplyr::rename(hr_emotion = Prediction, sr_emotion = Reference) %>% 
  dplyr::mutate(n_video = sum(Freq)) %>% 
  dplyr::mutate(prop_video = Freq/sum(Freq)) %>% 
  dplyr::group_by(sr_emotion) %>% 
  dplyr::mutate(n_sr = sum(Freq)) %>%
  dplyr::mutate(prop_sr = Freq/sum(Freq)) %>% 
  dplyr::ungroup()

@@

Each emotion label used to describe encoders’ self-reported subjective feeling (*i.e.*, the label rated with the highest value) is compared with the emotion labels which were rated with the highest score by human observers. Results of the confusion matrix show a low agreement between the emotion felt by the encoder during the elicitation and the emotion recognized by the human observers (Accuracy \nolinebreak = \nolinebreak `r round(confusionMatrix_sr_hr$overall["Accuracy"],2)`, 95% CI [`r round(confusionMatrix_sr_hr$overall["AccuracyLower"],2)`,`r round(confusionMatrix_sr_hr$overall["AccuracyUpper"],2)`]; Kappa = `r round(confusionMatrix_sr_hr$overall["Kappa"],2)`) except for *disgust* (`r scales::percent(as.numeric(confusionMatrix_sr_hr_freq[confusionMatrix_sr_hr_freq$hr_emotion == "disgust" & confusionMatrix_sr_hr_freq$sr_emotion == "disgust", "prop_sr"]))` of the videos self-reported). These results are far from those classically obtained in the literature for emotional facial expression recognition which ranges between 60% and 80% accuracy. However these results are mostly obtained with static (*i.e.*, pictures) and posed (*i.e.*, displayed by actors) facial expressions using only 6 emotional labels in a forced-choice paradigm.

@@

Sensitivity, specificity, precision and F1 scores for each emotion reveals that *happiness* has the highest coherence ratio whereas *sadness* has the lowest coherence ratio between true positives and false positives (Table \ref{table:confusionTable_sr_hr}).

```{r confusionTable_sr_hr, results="asis"}
confusionMatrix_sr_hr_table <- confusionMatrix_sr_hr %>%
  magrittr::use_series(byClass) %>%
  as.data.frame() %>%
  tibble::rownames_to_column("Emotion") %>%
  dplyr::mutate(Emotion = gsub(pattern = "Class: ",replacement = "",x = Emotion)) %>% 
  dplyr::select(Emotion, Sensitivity, Specificity, Precision, F1) %>% 
  dplyr::mutate_if(is.numeric,round,2) %>% 
  replace(is.na(.), "\\textit{na.}")

confusionMatrix_sr_hr_table %>%
  knitr::kable(
    "latex",
    caption = "\\label{table:confusionTable_sr_hr}Human recognition accuracy metrics for each emotion.",
    booktabs = TRUE,
    digits = 2,
    linesep = "",
    escape = FALSE,
    align = c('l',rep('c', 4))
  ) %>%
  kableExtra::kable_styling(
    latex_options = "HOLD_position",
    font_size = 8,
    full_width = TRUE 
    ) %>%
   add_footnote("Note. \\textit{na.} values are produced when not enough data are available to compute accuracy indicators.", notation="none", escape = FALSE)
```

@@

```{r sr_ar}
comparison_sr_ar <- dplyr::inner_join(self_report_score,automatic_recognition_score, by = "C_Video") %>% 
  dplyr::ungroup() %>% 
  dplyr::filter(ar_emotion != "undetermined") %>% 
  dplyr::filter(sr_emotion != "undetermined") %>% 
  dplyr::mutate_if(is.character,as.factor)

@@

Similarly to the previous analysis, the overall correlation of recognition and non-recognition revealed a significant but very low coherence between self-reported emotions and automatic classifier’s recognition (`r corr_sr_ar$full_result`). A confusion matrix was used to compare encoders’ subjective feeling with the emotion label recognized by the automatic classifier (Figure \ref{fig:confusionMatrix_sr_ar}).

```{r confusionMatrix_sr_ar, fig.height=6, fig.cap="\\label{fig:confusionMatrix_sr_ar}Confusion matrix of between the emotion self-reported as being characteristic of the elicitation with the emotion recognized by the automatic classifier."}
confusionMatrix_sr_ar_freq <- confusionMatrix_sr_ar$table %>% 
  as.data.frame() %>% 
  dplyr::rename(ar_emotion = Prediction, sr_emotion = Reference) %>% 
  dplyr::mutate(n_video = sum(Freq)) %>% 
  dplyr::mutate(prop_video = Freq/sum(Freq)) %>% 
  dplyr::group_by(sr_emotion) %>% 
  dplyr::mutate(n_sr = sum(Freq)) %>%
  dplyr::mutate(prop_sr = Freq/sum(Freq)) %>% 
  dplyr::ungroup()

@@

Results obtained for the comparison between emotions self-reported and recognized by the automatic classifier are somewhat similar to the ones with human observers (Table \ref{table:confusionTable_sr_ar}). Overall, a low agreement between emotion self-reported and emotion recognized by the automatic classifier (Accuracy = `r round(confusionMatrix_sr_ar$overall["Accuracy"],2)`, 95% CI [`r round(confusionMatrix_sr_ar$overall["AccuracyLower"],2)`,`r round(confusionMatrix_sr_ar$overall["AccuracyUpper"],2)`]; Kappa = `r round(confusionMatrix_sr_ar$overall["Kappa"],2)`) except for *happiness* (`r scales::percent(as.numeric(confusionMatrix_sr_ar_freq[confusionMatrix_sr_ar_freq$ar_emotion == "happiness" & confusionMatrix_sr_ar_freq$sr_emotion == "happiness", "prop_sr"]))` of the video self-reported) is evident.

```{r confusionTable_sr_ar, results="asis"}
confusionMatrix_sr_ar_table <- confusionMatrix_sr_ar %>%
  magrittr::use_series(byClass) %>%
  as.data.frame() %>%
  tibble::rownames_to_column("Emotion") %>%
  dplyr::mutate(Emotion = gsub(pattern = "Class: ",replacement = "",x = Emotion)) %>% 
  dplyr::select(Emotion, Sensitivity, Specificity, Precision, F1) %>% 
  dplyr::mutate_if(is.numeric,round,2) %>% 
  replace(is.na(.), "\\textit{na.}")

confusionMatrix_sr_ar_table %>%
  knitr::kable(
    "latex",
    caption = "\\label{table:confusionTable_sr_ar}Autonatic recognition accuracy metrics for each emotion.",
    booktabs = T,
    digits = 2,
    linesep = "",
    escape = FALSE,
    align = c('l',rep('c', 4))
  ) %>%
  kableExtra::kable_styling(
    latex_options = c("HOLD_position"),
    font_size = 8,
    full_width = TRUE
    ) %>%
   add_footnote("Note. \\textit{na.} values are produced when not enough data are available to compute accuracy indicators.", notation="none", escape = FALSE)
```

@@

A third confusion matrix is used to compare similarities (diagonal) and differences between human observers and automatic classifier in classifying the six emotion labels (Figure \ref{fig:confusionMatrix_hr_ar}).

```{r confusionMatrix_hr_ar, fig.height=6, fig.cap="\\label{fig:confusionMatrix_hr_ar}Proportion of emotion labels classified by human observers which are recognized by the automatif classifier."}
comparison_hr_ar <- dplyr::inner_join(human_recognition_score,automatic_recognition_score, by = "C_Video") %>% 
  dplyr::ungroup() %>% 
  dplyr::filter(ar_emotion != "undetermined") %>% 
  dplyr::filter(hr_emotion != "undetermined") %>% 
  dplyr::mutate_if(is.character,as.factor)

@@

Despite being one of the most investigated questions in affective science, the consistency between emotion felt and facially displayed on the one hand, and facial expression recognized on the other is a hot topic. To date no clear evidence has been found to definitively solve the questions raised. Yet, with the growing interest of industries and government in monitoring individual’s psychological states, this issue is under intense scrutiny. The present research aims to provide some empirical data to answer some of the questions posed. The faces encoders spontaneously displayed when confronted with an emotional eliciting task were submitted both to human and to automatic recognition. The criterion for recognition accuracy was the subjective feeling self-reported by the encoder once the elicitation task was carried out. Results first reveal a low consistency between emotion felt and facial expression displayed. They show that facial expressions of emotion are often not displayed when the Basic Emotion View would predict them to be expressed. Secondly, results show low accuracy rates for both humans and the automatic classifier in identifying the inner emotional states of these encoders based on their facial expressions. Thirdly, human observers prove to be better at recognizing the emotion facially expressed than the automatic recognition tool is. Such results support the hypothesis advanced by some authors of low emotion–expression coherence (Kappas 2003). In many instances, facial displays are not associated with a concordant emotional state, even any emotional state at all (Bonanno & Keltner 2004; Fernández-Dols & Crivelli, 2013]. More and more evidence is showing that facial expressions are in reality not expressing emotions (McKeown 2013]. Increasing studies show that, for most emotions, the EFE elicited by emotional triggers are scarce and partial, even when micro-expressions are taken into account (Dúran et al., 2017). These studies are conducted either in laboratory settings or in the field. This is not to say that facial expression is not an informative modality for understanding the emotional state of a person. There is indeed an affinity between emotion and facial display (Frijda & Tcherkassof, 1997). However, claiming that there are unique “signatures” allowing using configurations of facial muscles to infer the presence of a specific emotion is misleading. As well as other nonverbal behaviors, facial movements are not only assumed to be determined by emotion but also by various other causes, such as psychological states (e.g., motivations or pain), to say nothing of social context and sociocultural norms (Ekman et al. 1987). Hence, facial movements have causes and functions other than the expression of emotion. This multiple determination excludes any possibility of drawing a linear inference from facial activity on the underlying psychological state (emotional or other). Beyond the present observations showing a weak consistency between subjective feelings and spontaneous facial expressions, this study sheds some light on the controversy between the Basic Emotion View and the constructivist approach as to the facial recognition issue. The former (Ekman 1992) assumes that expressions of emotion are brief and coherent patterns of facial muscle movements that co-vary with discrete subjective experiences. In return, this information displayed by the face corresponds to the one extracted by perceivers. Instead of viewing emotions as natural kinds (Barrett, 2006), the constructivist approach supposes that emotions are social constructions and that facial behaviors intrinsically situated. The emotions that are recognized by the observer are constructed in her/his mind. Therefore, facial movements do not express specific emotions because they do not carry intrinsic emotional signification. It is the observer that infers the emotional meaning of the facial expression, this interpretation depending on availability of different kinds of information (contextual information, linguistic categories…). As a consequence, one can predict from the first line of thinking that individuals’ emotional subjective feeling should be correlated to the recognition of facial expressions from both human observers and automatic classifiers whereas if emotions are social constructs, as stated by the second line of thinking, human observer’s should be better at perceiving emotions expressed on the face than automatic classifiers.

>>>>

keywords          : "Facial expression, self-report, human observer, automatic recognition."
wordcount         : "7998"

@@

header-includes:
  - \usepackage{booktabs}
  - \usepackage{float}
  - \usepackage{tabu}
  - \usepackage{wrapfig}
  - \usepackage[none]{hyphenat}
  - \floatplacement{figure}{H}
  - \usepackage{flushend}
  - \usepackage{biblatex}
---
(ref:confusionMatrix-sr-hr) Confusion matrix between the emotion self-reported as being characteristic of the elicitation with the emotion recognized by the human observers.
(ref:confusionMatrix-sr-ar) Confusion matrix of between the emotion self-reported as being characteristic of the elicitation with the emotion recognized by the automatic classifier.
(ref:confusionMatrix-hr-ar) Proportion of emotion labels classified by human observers which are recognized by the automatif classifier.
(ref:confusionTable-sr-hr) Human recognition accuracy metrics for each emotion.
(ref:confusionTable-sr-ar) Autonatic recognition accuracy metrics for each emotion.

@@

With the development of commercial automatic facial expression recognition tools [@dupre2018accuracy], industries and governments are gradually implementing this technology in order to monitor humans’ emotions in various scenarios (*e.g.*, in marketing, healthcare, and the automotive industry to name a few). This technology rests on the premise that facial expressions provide a direct access to individuals’ subjective feelings, and that one can read the emotions displayed on a person’s face as in an open book. This premise is derived from the Basic Emotion View which Paul Ekman is one of the famous proponents of this perspective [@ekman1987universals; @ekman1988universality; @ekman1992argument; @ekman2007directed]. Even if this premise is central to the modern mainstream approach of human emotion, recent research in affective science is challenging it. Once the Basic Emotion View is briefly described, its foremost criticisms are synthetically exposed. Inconsistencies and unpredicted findings research has uncovered come both from field observations and laboratory experiments. They relate to two different issues. The first issue is the one of the sender’s production of spontaneous facial displays. The second issue has to do with the viewer’s interpretation of these facial displays. Building on this evidence, and as a further attempt to shed light on the emotion–face link, the present study is finally exposed. Its main goal is to tackle these two issues at the same time. It aims at identifying the extent to which ordinary people’s experienced emotions are displayed through identifiable emotional facial expressions (EFE) recognized as such both by human individuals and by automatic facial expression recognition tools.

@@

```{r self-report-method}
self_report_gender <- self_report_data %>% 
  dplyr::group_by(genre_c) %>% 
  dplyr::summarise(n = n())

@@

```{r human-recognition-method}
unique_ppt_count <- human_recognition_data %>% 
  dplyr::group_by(source,SEXE_Juge, C_Juge) %>% 
  dplyr::summarise(n = n())

@@

```{r sr-hr}
comparison_sr_hr <- dplyr::inner_join(self_report_score,human_recognition_score, by = "C_Video") %>% 
  dplyr::ungroup() %>% 
  dplyr::filter(hr_emotion != "undetermined") %>% 
  dplyr::filter(sr_emotion != "undetermined") %>% 
  dplyr::mutate_if(is.character,as.factor)

@@

The overall correlation of recognition and non-recognition between self-reported emotions and human observers recognition is significant but low (`r corr_sr_hr$full_result`). In order to identify differences according to the emotional labels, encoders’ subjective feelings are compared with human observers’ recognition in a confusion matrix (Figure \@ref(fig:confusionMatrix-sr-hr).

```{r confusionMatrix-sr-hr, fig.height=6, out.width ="\\textwidth", fig.pos = "!h", fig.cap="(ref:confusionMatrix-sr-hr)"}
confusionMatrix_sr_hr_freq <- confusionMatrix_sr_hr$table %>% 
  as.data.frame() %>% 
  dplyr::rename(hr_emotion = Prediction, sr_emotion = Reference) %>% 
  dplyr::mutate(n_video = sum(Freq)) %>% 
  dplyr::mutate(prop_video = Freq/sum(Freq)) %>% 
  dplyr::group_by(sr_emotion) %>% 
  dplyr::mutate(n_sr = sum(Freq)) %>%
  dplyr::mutate(prop_sr = Freq/sum(Freq)) %>% 
  dplyr::ungroup()

@@

Each emotion label used to describe encoders’ self-reported subjective feeling (*i.e.*, the label rated with the highest value) is compared with the emotion labels which were rated with the highest score by human observers. Results of the confusion matrix show a low agreement between the emotion felt by the encoder during the elicitation and the emotion recognized by the human observers (Accuracy = `r round(confusionMatrix_sr_hr$overall["Accuracy"],2)`, 95% CI [`r round(confusionMatrix_sr_hr$overall["AccuracyLower"],2)`,`r round(confusionMatrix_sr_hr$overall["AccuracyUpper"],2)`]; Kappa = `r round(confusionMatrix_sr_hr$overall["Kappa"],2)`) except for *disgust* (`r scales::percent(as.numeric(confusionMatrix_sr_hr_freq[confusionMatrix_sr_hr_freq$hr_emotion == "disgust" & confusionMatrix_sr_hr_freq$sr_emotion == "disgust", "prop_sr"]))` of the videos self-reported). These results are far from those classically obtained in the literature for emotional facial expression recognition which ranges between 60% and 80% accuracy. However these results are mostly obtained with static (*i.e.*, pictures) and posed (*i.e.*, displayed by actors) facial expressions using only 6 emotional labels in a forced-choice paradigm.

@@

Sensitivity, specificity, precision and F1 scores for each emotion reveals that *happiness* has the highest coherence ratio whereas *sadness* has the lowest coherence ratio between true positives and false positives (Table \@ref(tab:confusionTable-sr-hr)).

```{r confusionTable-sr-hr, results="asis"}
confusionMatrix_sr_hr_table <- confusionMatrix_sr_hr %>%
  magrittr::use_series(byClass) %>%
  as.data.frame() %>%
  tibble::rownames_to_column("Emotion") %>%
  dplyr::mutate(Emotion = gsub(pattern = "Class: ",replacement = "",x = Emotion)) %>% 
  dplyr::select(Emotion, Sensitivity, Specificity, Precision, F1) %>% 
  dplyr::mutate_if(is.numeric,round,2) %>% 
  replace(is.na(.), "\\textit{na.}")

confusionMatrix_sr_hr_table %>%
  knitr::kable(
    "latex",
    caption = "(ref:confusionTable-sr-hr)",
    booktabs = TRUE,
    digits = 2,
    linesep = "",
    escape = FALSE,
    align = c('l', rep('c', 4)),
    longtable = T
  ) %>%
  kableExtra::kable_styling(
    latex_options = "HOLD_position",
    font_size = 8,
    full_width = TRUE 
  ) %>%
  kableExtra::footnote(
    general_title = "",
    general = "Note. *na.* values are produced when not enough data are available to compute accuracy indicators.", 
    escape = FALSE,
    footnote_as_chunk = TRUE,
    threeparttable = TRUE
  )
```

@@

```{r sr-ar}
comparison_sr_ar <- dplyr::inner_join(self_report_score,automatic_recognition_score, by = "C_Video") %>% 
  dplyr::ungroup() %>% 
  dplyr::filter(ar_emotion != "undetermined") %>% 
  dplyr::filter(sr_emotion != "undetermined") %>% 
  dplyr::mutate_if(is.character,as.factor)

@@

Similarly to the previous analysis, the overall correlation of recognition and non-recognition revealed a significant but very low coherence between self-reported emotions and automatic classifier’s recognition (`r corr_sr_ar$full_result`). A confusion matrix was used to compare encoders’ subjective feeling with the emotion label recognized by the automatic classifier (Figure \@ref(fig:confusionMatrix-sr-ar)).

```{r confusionMatrix-sr-ar, fig.height=6, out.width ="\\textwidth", fig.pos = "!h", fig.cap="(ref:confusionMatrix-sr-ar)"}
confusionMatrix_sr_ar_freq <- confusionMatrix_sr_ar$table %>% 
  as.data.frame() %>% 
  dplyr::rename(ar_emotion = Prediction, sr_emotion = Reference) %>% 
  dplyr::mutate(n_video = sum(Freq)) %>% 
  dplyr::mutate(prop_video = Freq/sum(Freq)) %>% 
  dplyr::group_by(sr_emotion) %>% 
  dplyr::mutate(n_sr = sum(Freq)) %>%
  dplyr::mutate(prop_sr = Freq/sum(Freq)) %>% 
  dplyr::ungroup()

@@

Results obtained for the comparison between emotions self-reported and recognized by the automatic classifier are somewhat similar to the ones with human observers (Table \@ref(tab:confusionTable-sr-ar)). Overall, a low agreement between emotion self-reported and emotion recognized by the automatic classifier (Accuracy = `r round(confusionMatrix_sr_ar$overall["Accuracy"],2)`, 95% CI [`r round(confusionMatrix_sr_ar$overall["AccuracyLower"],2)`,`r round(confusionMatrix_sr_ar$overall["AccuracyUpper"],2)`]; Kappa = `r round(confusionMatrix_sr_ar$overall["Kappa"],2)`) except for *happiness* (`r scales::percent(as.numeric(confusionMatrix_sr_ar_freq[confusionMatrix_sr_ar_freq$ar_emotion == "happiness" & confusionMatrix_sr_ar_freq$sr_emotion == "happiness", "prop_sr"]))` of the video self-reported) is evident.

```{r confusionTable-sr-ar, results="asis"}
confusionMatrix_sr_ar_table <- confusionMatrix_sr_ar %>%
  magrittr::use_series(byClass) %>%
  as.data.frame() %>%
  tibble::rownames_to_column("Emotion") %>%
  dplyr::mutate(Emotion = gsub(pattern = "Class: ",replacement = "",x = Emotion)) %>% 
  dplyr::select(Emotion, Sensitivity, Specificity, Precision, F1) %>% 
  dplyr::mutate_if(is.numeric,round,2) %>% 
  replace(is.na(.), "\\textit{na.}")

confusionMatrix_sr_ar_table %>%
  knitr::kable(
    "latex",
    caption = "(ref:confusionTable-sr-ar)",
    booktabs = T,
    digits = 2,
    linesep = "",
    escape = FALSE,
    align = c('l',rep('c', 4)),
    longtable = T
  ) %>%
  kableExtra::kable_styling(
    latex_options = c("HOLD_position"),
    font_size = 8,
    full_width = TRUE
    ) %>%
  kableExtra::footnote(
    general_title = "",
    general = "Note. *na.* values are produced when not enough data are available to compute accuracy indicators.", 
    escape = FALSE,
    footnote_as_chunk = TRUE,
    threeparttable = TRUE
    )
```

@@

A third confusion matrix is used to compare similarities (diagonal) and differences between human observers and automatic classifier in classifying the six emotion labels (Figure \@ref(fig:confusionMatrix-hr-ar)).

```{r confusionMatrix-hr-ar, fig.height=6, out.width ="\\textwidth", fig.pos = "!h", fig.cap="(ref:confusionMatrix-hr-ar)"}
comparison_hr_ar <- dplyr::inner_join(human_recognition_score,automatic_recognition_score, by = "C_Video") %>% 
  dplyr::ungroup() %>% 
  dplyr::filter(ar_emotion != "undetermined") %>% 
  dplyr::filter(hr_emotion != "undetermined") %>% 
  dplyr::mutate_if(is.character, as.factor)

@@

Despite being one of the most investigated questions in affective science, the consistency between emotion felt and facially displayed on the one hand, and facial expression recognized on the other is a hot topic. To date no clear evidence has been found to definitively solve the questions raised. Yet, with the growing interest of industries and government in monitoring individual’s psychological states, this issue is under intense scrutiny. The present research aims to provide some empirical data to answer some of the questions posed. The faces encoders spontaneously displayed when confronted with an emotional eliciting task were submitted both to human and to automatic recognition. The criterion for recognition accuracy was the subjective feeling self-reported by the encoder once the elicitation task was carried out. Results first reveal a low consistency between emotion felt and facial expression displayed. They show that facial expressions of emotion are often not displayed when the Basic Emotion View would predict them to be expressed. Secondly, results show low accuracy rates for both humans and the automatic classifier in identifying the inner emotional states of these encoders based on their facial expressions. Thirdly, human observers prove to be better at recognizing the emotion facially expressed than the automatic recognition tool is. Such results support the hypothesis advanced by some authors of low emotion–expression coherence [@kappas2003facial]. In many instances, facial displays are not associated with a concordant emotional state, even any emotional state at all [@bonanno2004brief; @fernandez2013emotion]. More and more evidence is showing that facial expressions are in reality not expressing emotions [@mckeown2013analogical]. Increasing studies show that, for most emotions, the EFE elicited by emotional triggers are scarce and partial, even when micro-expressions are taken into account [@duran2017coherence]. These studies are conducted either in laboratory settings or in the field. This is not to say that facial expression is not an informative modality for understanding the emotional state of a person. There is indeed an affinity between emotion and facial display [@frijda1997facial]. However, claiming that there are unique “signatures” allowing using configurations of facial muscles to infer the presence of a specific emotion is misleading. As well as other nonverbal behaviors, facial movements are not only assumed to be determined by emotion but also by various other causes, such as psychological states (*e.g.*, motivations or pain), to say nothing of social context and sociocultural norms [@ekman1987universals]. Hence, facial movements have causes and functions other than the expression of emotion. This multiple determination excludes any possibility of drawing a linear inference from facial activity on the underlying psychological state (emotional or other). Beyond the present observations showing a weak consistency between subjective feelings and spontaneous facial expressions, this study sheds some light on the controversy between the Basic Emotion View and the constructivist approach as to the facial recognition issue. The former [@ekman1992argument] assumes that expressions of emotion are brief and coherent patterns of facial muscle movements that co-vary with discrete subjective experiences. In return, this information displayed by the face corresponds to the one extracted by perceivers. Instead of viewing emotions as natural kinds [@barrett2006emotions], the constructivist approach supposes that emotions are social constructions and that facial behaviors intrinsically situated. The emotions that are recognized by the observer are constructed in her/his mind. Therefore, facial movements do not express specific emotions because they do not carry intrinsic emotional signification. It is the observer that infers the emotional meaning of the facial expression, this interpretation depending on availability of different kinds of information (contextual information, linguistic categories...). As a consequence, one can predict from the first line of thinking that individuals’ emotional subjective feeling should be correlated to the recognition of facial expressions from both human observers and automatic classifiers whereas if emotions are social constructs, as stated by the second line of thinking, human observer’s should be better at perceiving emotions expressed on the face than automatic classifiers.
